A Bayesian ranking statistic to find high-mass black holes in LIGO--Virgo data

### 2021-03-29 main.tex

%%%%%%---SECTIONS---%%%%%%%%%%%%
\section{Introduction}
In the 1970s, X-ray and optical observations from the Cyg-X1 binary star system led to the discovery of the first stellar-mass black hole~\cite{Webster:1972:Natur}. Two years later, in 1974, radio observation of our galaxy center demonstrated the presence of a supermassive black hole -- Sgr A$^*$~\cite{Balick:1974:ApJ}. In the half-century since then, astronomers accumulated more definitive evidence that have established the  existence of stellar-mass black holes $\text{M}_\text{BH} < 10^{2} \ \msun$ and supermassive black holes, $\text{M}_\text{BH} > 10^{5} \ \msun$ in our universe (e.g. ~\cite{Ghez:1998:ApJ, Genzel:2010:RvMP, Abbott:2019:PhRvX, EventHorizonTelescopeCollaboration:2019:ApJL, Abbott:2020:arXiv}). However, a gap in between the mass-ranges of these two categories has formed: there is a deficiency of observational evidence for black holes in the intermediate-mass range $10^{2} - 10^{5}\ \text{M}_{\odot}$. These intermediate-mass black holes must have existed to form the supermassive black holes detected to exist in the early universe~\cite{Banados:2018:Natur}. However, we only have a handful of promising candidates~\cite{Greene:2004:ApJ, Graham:2013:ApJ, Mezcua:2017:IJMPD, Koliopanos:2017:mbhe, Lin:2020:ApJL, Greene:2020:ARA&A}, and one confirmed intermediate-mass black hole, the remnant of the gravitational wave event GW190521 with a mass of $142^{+28}_{-16}\ \msun$~\cite{Abbott:2020:PhRvL}. The concrete discovery of more intermediate-mass black holes will bridge the observational gap and illuminate our understanding of galaxy and supermassive black hole formation.

@@

Hierarchical mergers in active galactic nuclei migration-traps, globular and stellar clusters may be a feasible light-seed method for black holes to accumulate mass and reach the intermediate-mass range. \rs{\sout{However, when black holes merge, they can receive recoil-kicks up to $\sim$2,500 km/s~\cite{}. Hence, the merger remnants require their recoil kick to be less than its environment's escape velocity to undergo successive mergers. This tight constraint makes hierarchical mergers a challenging scenario.}

@@

The standard framework to identify CBC gravitational-wave signals hidden in data is by quantifying the significance of candidates with null-hypothesis significance testing. In this framework, the candidates' ranking statistic is compared against a background distribution in a frequentist approach. On the other hand, the standard framework for performing parameter estimation and model selection in gravitational-wave astronomy is Bayesian inference. This work utilises Bayesian inference to calculate the Bayesian Coherence odds-ratio~\cite{BCR1}, \bcr, of high-mass candidates in LIGO's second observing run. We use the \bcr not as an odds-ratio but instead as a ranking statistic, a step toward building a unified Bayesian framework to search for candidates and estimate their parameters.

@@

\subsection{IAS Events}
Our analysis of the high-mass IAS events and candidates in O2 has resulted in three events with disfavourable $\pastrobcr<0.5$ (GWC170402, GW170403, GW170425), and four events and one candidate with $\pastrobcr\geq0.5$ (GW170121, 170302, GW170304, GW170727, GW170817A).

GWC170402, detected by \citet{IAS2}, is reported to have a signal that is not described well by waveforms for circular binaries with aligned spins. Hence, we might have received a low \pastrobcr due to our usage of \imrphenomp, a waveform that does not account for eccentricity. Additionally, the search conducted by \citet{IAS2} was a single-detector search. Our ranking statistic relies on the signal to appear coherent, even if just faintly coherent, amongst the various detectors to have a high \pastrobcr. The lack of coherence and the non-eccentric waveform may be the leading factors for a low \pastro. GW170403 and GW170425 which have $\pastrobcr<0.35$ also have low  \pastro reported by \citet{pycbc_ogc_2},  suggesting that these events may have been false alarms.

>>>>

%%%%%%---SECTIONS---%%%%%%%%%%%%
\section{Introduction}
Since the 1970s, there has been a steady accumulation of evidence for stellar mass ($\text{M}_\text{BH} < 10^{2} \ \msun$) and supermassive black holes ($\text{M}_\text{BH} > 10^{5} \ \msun$)~cite{Webster:1972:Natur, Balick:1974:ApJ, Ghez:1998:ApJ, Genzel:2010:RvMP, Abbott:2019:PhRvX, EventHorizonTelescopeCollaboration:2019:ApJL, Abbott:2020:arXiv}. However, there is a deficiency of observational evidence for black holes in the intermediate-mass range $10^{2} - 10^{5}\ \text{M}_{\odot}$. Given the existence of supermassive black holes detected in the early universe, it is likely for intermediate-mass black holes to have also existed~\cite{Banados:2018:Natur}. However, we only have a handful of promising candidates~\cite{Greene:2004:ApJ, Graham:2013:ApJ, Mezcua:2017:IJMPD, Koliopanos:2017:mbhe, Lin:2020:ApJL, Greene:2020:ARA&A}, and one confirmed intermediate-mass black hole, the remnant of the gravitational wave event GW190521 with a mass of $142^{+28}_{-16}\ \msun$~\cite{Abbott:2020:PhRvL}. The concrete discovery of more intermediate-mass black holes will bridge the observational gap and illuminate our understanding of galaxy and supermassive black hole formation.

@@

Hierarchical mergers in active galactic nuclei migration-traps, globular and stellar clusters may be a feasible light-seed method for black holes to accumulate mass and reach the intermediate-mass range. \rs{\sout{However, when black holes merge, they can receive recoil-kicks up to $\sim$2,500 km/s~\cite{}. Hence, the merger remnants require their recoil kick to be less than its environment's escape velocity to undergo successive mergers. This tight constraint makes hierarchical mergers a challenging scenario.}}

@@

The standard framework to identify CBC gravitational-wave signals hidden in data is by quantifying the significance of candidates with null-hypothesis significance testing. In this framework, the candidates' ranking statistic is compared against a background distribution in a frequentist approach. \rs{add a sentence describing what the ranking stats are for the different pipelines (key words: matched filtering, SNR etc...) and add citations!} On the other hand, the standard framework for performing parameter estimation and model selection in gravitational-wave astronomy is Bayesian inference. This work utilises Bayesian inference to calculate the Bayesian Coherence odds-ratio~\cite{BCR1}, \bcr, of high-mass candidates in LIGO's second observing run. We use the \bcr not as an odds-ratio but instead as a ranking statistic, a step toward building a unified Bayesian framework to search for candidates and estimate their parameters.

@@

\subsection{IAS Events}
Our analysis of the high-mass IAS events and candidates in O2 has resulted in three events with disfavourable $\pastrobcr<0.5$ (GWC170402, GW170403, GW170425), and four events and one candidate with $\pastrobcr\geq0.5$ (GW170121, 170302, GW170304, GW170727, GW170817A). \rs{suggest an additional sentence comparing the numbers with greater/less than 0.5 to the numbers that IAS find.}

GWC170402, detected by \citet{IAS2}, is reported to have a signal that is not described well by waveforms for circular binaries with aligned spins \rs{what does "not well described" mean?}. Hence, we might have received a low \pastrobcr due to our usage of \imrphenomp, a waveform that does not account for eccentricity. Additionally, the search conducted by \citet{IAS2} was a single-detector search. Our ranking statistic relies on the signal to appear coherent, even if just faintly coherent, amongst the various detectors to have a high \pastrobcr. The lack of coherence and the non-eccentric waveform may be the leading factors for a low \pastro. GW170403 and GW170425 which have $\pastrobcr<0.35$ also have low  \pastro reported by \citet{pycbc_ogc_2},  suggesting that these events may have been false alarms.
